---
title: "Using Scales to Customize Plots"
output: learnr::tutorial
runtime: shiny_prerendered
---

```{r setup, include=FALSE}
library(learnr)
library(tidyverse)
library(medicaldata)
library(janitor)
library(scales)
library(nycflights13)
knitr::opts_chunk$set(echo = FALSE)
prostate <- medicaldata::blood_storage %>% 
  clean_names() %>% as_tibble()
covid <- medicaldata::covid_testing %>% clean_names() %>%  as.tibble()
laryngoscope <- medicaldata::laryngoscope %>% clean_names() %>%  as.tibble()
licorice <- medicaldata::licorice_gargle %>% clean_names() %>%  as.tibble()
```


## Customizing X and Y axes

### Exercise 1

The R code below will plot the bmi (x axis) against the total_intubation_time (y axis) as a scatterplot.
Your challenge:
- Make the Y axis start at exactly 0 (turn off multiplicative expansion) and end at 105
- make the X axis breaks at every 5 units of BMI from 30 to 65

```{r axes, exercise=TRUE, error=TRUE, lines = 5}
laryngoscope %>% 
  ggplot(aes(x= bmi, y = total_intubation_time)) + 
  geom_point()
```

```{r axes-hint, error=TRUE}
laryngoscope %>% 
  ggplot(aes(x= bmi, y = total_intubation_time)) + 
  geom_point() +
  scale_y_continuous(limits = ---,
                     expand = ---) +
  scale_x_continuous(breaks = seq(---))
```

```{r axes-solution}
laryngoscope %>% 
  ggplot(aes(x= bmi, y = total_intubation_time)) + 
  geom_point() +
  scale_y_continuous(limits = c(0,105),
        expand = expansion(mult = 0)) +
  scale_x_continuous(breaks = seq(from = 30, to = 65, by=5))
```
### Exercise 2

You can use the `expand` argument to control the margins of a plot. These can be expanded additively or multiplicatively. The default for discrete scales is to add 0.6 of a category on each side. The default for continuous scales is to multiply by 5% of the range on each side.

In this exercise, try to
- Make more room at the top for the legend (multiply by 20% of y on top only - leave bottom at 5%)
- rename the legend title to "View Achieved"


```{r scale2, exercise=TRUE, error=TRUE, lines = 5}
laryngoscope %>% 
  ggplot(aes(x= asa, y = total_intubation_time, color = factor(view))) + 
  geom_jitter() +
  theme(legend.position = c(0.5, 0.9)) +
  scale_color_discrete(guide = guide_legend(nrow = 1))
  
```

```{r scale2-hint}
laryngoscope %>% 
  ggplot(aes(x= asa, y = total_intubation_time, color = factor(view))) + 
  geom_jitter() +
  theme(legend.position = c(0.5, 0.9)) +
  scale_color_discrete(guide = guide_legend(nrow = 1),
    name = ---) +
  scale_y_continuous(expand = expansion(add = c(--,--)))

```

```{r scale2-solution}
laryngoscope %>% 
  ggplot(aes(x= asa, y = total_intubation_time, color = factor(view))) + 
  geom_jitter() +
  theme(legend.position = c(0.5, 0.9)) +
  scale_color_discrete(guide = guide_legend(nrow = 1),
    name = "View Achieved") +
  scale_y_continuous(expand = expansion(mult = c(0.05,0.2)))

```

### Exercise 3

You can customize the appearance of dates on scales, with `scale_(x|y)_date()` functions.

The code below will plot the average NYC flight delay by date in 2013. Your challenge is to:
- Set the x axis breaks to monthly
- format the labels to abbreviated month and 2-digit year, with a 2-line format, using `\n`

```{r scale3, exercise=TRUE, error=TRUE, lines = 5}
nycflights13::flights %>% 
  mutate(date= lubridate::dmy(paste(day, month, year, sep = "-"))) %>% 
  group_by(date) %>% 
  summarize(mean_delay_min = mean(arr_delay, na.rm = TRUE)) %>% 
  ggplot(aes(x = date, y = mean_delay_min)) +
  geom_point() +
  scale_x_date(---)
```

```{r scale3-hint}
nycflights13::flights %>% 
  mutate(date= lubridate::dmy(paste(day, month, year, sep = "-"))) %>% 
  group_by(date) %>% 
  summarize(mean_delay_min = mean(arr_delay, na.rm = TRUE)) %>% 
  ggplot(aes(x = date, y = mean_delay_min)) +
  geom_point() +
  scale_x_date(date_breaks = "---",
               date_labels = "---")
```

```{r scale3-solution}
nycflights13::flights %>% 
  mutate(date= lubridate::dmy(paste(day, month, year, sep = "-"))) %>% 
  group_by(date) %>% 
  summarize(mean_delay_min = mean(arr_delay, na.rm = TRUE)) %>% 
  ggplot(aes(x = date, y = mean_delay_min)) +
  geom_point() +
  scale_x_date(date_breaks = "1 month",
               date_labels = "%b\n%y")
```
### Exercise 4

Edit the R code below to format the previous plot with breaks every 4 weeks, with labels using the digit week of the year (starting on Sunday).

```{r scale4, exercise=TRUE, error=TRUE, lines = 5}
nycflights13::flights %>% 
  mutate(date= lubridate::dmy(paste(day, month, year, sep = "-"))) %>% 
  group_by(date) %>% 
  summarize(mean_delay_min = mean(arr_delay, na.rm = TRUE)) %>% 
  ggplot(aes(x = date, y = mean_delay_min)) +
  geom_point() 
```

```{r scale4-hint, error=TRUE, lines = 5}
nycflights13::flights %>% 
  mutate(date= lubridate::dmy(paste(day, month, year, sep = "-"))) %>% 
  group_by(date) %>% 
  summarize(mean_delay_min = mean(arr_delay, na.rm = TRUE)) %>% 
  ggplot(aes(x = date, y = mean_delay_min)) +
  geom_point() +
  scale_x_date(date_breaks = "--",
               date_labels = "--")
```

```{r scale4-solution}
nycflights13::flights %>% 
  mutate(date= lubridate::dmy(paste(day, month, year, sep = "-"))) %>% 
  group_by(date) %>% 
  summarize(mean_delay_min = mean(arr_delay, na.rm = TRUE)) %>% 
  ggplot(aes(x = date, y = mean_delay_min)) +
  geom_point() +
  scale_x_date(date_breaks = "4 week",
               date_labels = "%U")
```


