---
title: "Counting and Tallying Variables"
output: learnr::tutorial
runtime: shiny_prerendered
---

```{r setup, include=FALSE}
library(learnr)
library(tidyverse)
library(medicaldata)
library(janitor)
library(gradethis)
knitr::opts_chunk$set(echo = FALSE)
prostate <- medicaldata::blood_storage %>% 
  clean_names() %>% as_tibble()
covid <- medicaldata::covid_testing %>% clean_names() %>%  as.tibble()
laryngoscope <- medicaldata::laryngoscope %>% clean_names() %>%  as.tibble()
licorice <- medicaldata::licorice_gargle %>% clean_names() %>%  as.tibble()
```


## Counts and Tallies of Variables in a dataset

Sometimes it is convenient to to have a total `tally()` (denominator) of your dataset, or by current groups. This helps in calculating proportions or percentages. You can also perform `counts` by groups and subgroups. The simplest way to count is with the `n()` function, which counts the observations and places the result in a vector named (by default) "n".

We will start with the laryngoscope dataset, seen here with `glimpse()`:

```{r , echo=TRUE}
glimpse(laryngoscope)
```

### Exercise 1

Write the R code required to tally the total number of observations with the`tally()` function, and name the new variable "observations".

```{r tally, exercise=TRUE, error=TRUE, lines = 5}
laryngoscope %>% 
  ---(name = "---")
```

```{r tally-hint}
laryngoscope %>% 
  tally(name = "observations")
```

### Exercise 2

If your data are currently grouped, `tally()` will count observations by group.

Write the R code required to (1) `group_by()` ASA class ("asa"), then (2) tally the total number of observations in each group with the`tally()` function, and name the new variable "obs".

```{r tally2, exercise=TRUE, error=TRUE, lines = 5}
laryngoscope %>% 
  ---(asa) %>% 
  ---(name = ---)
  
```

```{r tally2-hint}
laryngoscope %>% 
  group_by(asa) %>% 
  ---(name = ---)

```

```{r tally2-solution}
laryngoscope %>% 
  group_by(asa) %>% 
  tally(name = "obs")

```

### Exercise 3

You can do the grouping and counting in one function with `count()`, which will also ungroup after counting, so as not to change any previous grouping.

Write the R code required to (1) `group_by()` gender, then (2) count the total number of observations in each group with the`count()` function.

```{r count1, exercise=TRUE, error=TRUE, lines = 5}
laryngoscope %>% 
  count(---)
```

```{r count1-hint}
laryngoscope %>% 
  count(gender)
```

### Exercise 4

Write the R code required to count the total number of observations in each group of asa and gender with the`count()` function, then sort them in descending order, and name the count variable "n_per_group".

```{r count2, exercise=TRUE, error=TRUE, lines = 5}
laryngoscope %>% 
  count(---, ---, sort = ___, name = ___)
```

```{r count2-hint, error=TRUE, lines = 5}
laryngoscope %>% 
  count(asa, gender, sort = ___, name = ___)
```

```{r count2-solution}
laryngoscope %>% 
  count(asa, gender, sort = TRUE, name = "n_per_group")


```


### Exercise 5

Write the R code required to (1) add a "total_obs" column with the total number of observations in the study, and (2) add a "group_count" variable for each each combination of asa and gender, then (3) select asa, gender, group_count, and total_obs.

```{r count3, exercise=TRUE, error=TRUE, lines = 5}
laryngoscope %>% 
  add_tally(name = "---") %>% 
  add_count(--, --, -- = "---") %>% 
  select(---)
```

```{r count3-hint, error=TRUE, lines = 5}
laryngoscope %>% 
  add_tally(name = "---") %>% 
  add_count(asa, gender, name = "---") %>% 
  select(asa, gender, group_count, total_obs)
```

```{r count3-solution}
laryngoscope %>% 
  add_tally(name = "total_obs") %>% 
  add_count(asa, gender, name = "group_count") %>% 
  select(asa, gender, group_count, total_obs)
```

### Exercise 6

At times you just want a sense of how many values are in a particular character or factor variable. This is where the `janitor::tabyl()` function shines.
Write the R code required to count the number of distinct categories in the covid dataset variable, payor_group.

```{r tabyl1, exercise=TRUE, error=TRUE, lines = 5}
covid %>% 
  janitor::tabyl(---)
```

```{r tabyl1-hint, error=TRUE, lines = 5}
covid %>% 
  tabyl(---))
```

```{r tabyl1-solution}
covid %>% 
  tabyl(payor_group)
```

### Exercise 7

Write the R code required to count the number of observations by group in an R x C table in the covid dataset using the variables payor_group and gender.

```{r tabyl2, exercise=TRUE, error=TRUE, lines = 5}
covid %>% 
  ---(---, ---)
```

```{r tabyl2-hint, error=TRUE, lines = 5}
covid %>% 
  tabyl(---, ---)
```

```{r tabyl2-solution}
covid %>% 
  tabyl(payor_group, gender)
```
