---
title: "Tutorial"
output: learnr::tutorial
runtime: shiny_prerendered
---

```{r setup, include=FALSE}
library(learnr)
library(tidyverse)
library(medicaldata)
library(janitor)
knitr::opts_chunk$set(echo = FALSE)
prostate <- medicaldata::blood_storage %>% 
  clean_names() %>% as_tibble()
```


## Filtering Numeric Variables

We will start with the prostate dataset, seen here:

```{r echo=FALSE}
glimpse(prostate)
```

### Exercise 1

Write the R code required to filter the prostate dataset to rows with a prostate volume (p_vol) greater than or equal to 90:

```{r p_vol, exercise=TRUE, error=TRUE, lines = 5}
prostate %>% 
  filter(-- >= --)
```

```{r p_vol-hint}
prostate %>% 
  filter(p_vol >= 90)

```

### Exercise 2

Write the R code required to filter the prostate dataset to rows with a family history (fam_hx) of prostate cancer.

**Watch** the number of == signs

```{r fam_hx, exercise=TRUE, error=TRUE, lines = 5}
prostate %>% 
  select(age, t_vol, fam_hx) %>% 
  filter(fam_hx --)
```

```{r fam_hx-hint}
prostate %>% 
  select(age, t_vol, fam_hx) %>% 
  filter(fam_hx == 1)

```

### Exercise 3

Write the R code required to filter the prostate dataset to rows with a preoperative psa (preop_psa) near 12 (within 1).


```{r psa, exercise=TRUE, error=TRUE, lines = 5}
prostate %>% 
  select(age, aa, preop_psa) %>% 
  filter(preop_psa)
```

```{r psa-hint}
prostate %>% 
  select(age, aa, preop_psa) %>% 
  filter(near(preop_psa, 12, tol = 1))
```

### Exercise 4

Write the R code required to filter the prostate dataset to rows with ages with values of 60 or 63 or 69.

```{r in, exercise=TRUE, error=TRUE, lines = 5}
prostate %>% 
  select(age, preop_psa, fam_hx) %>% 
  filter()
```

```{r in-hint}
prostate %>% 
  select(age, preop_psa, fam_hx) %>% 
  filter(age %in% c(60, 63, 69))
```

### Exercise 5

Write the R code required to filter the prostate dataset to rows with preop_psa between 9 and 11.

```{r btwn, exercise=TRUE, error=TRUE, lines = 5}
prostate %>% 
  select(age, preop_psa, fam_hx) %>% 
  filter()
```

```{r btwn-hint}
prostate %>% 
  select(age, preop_psa, fam_hx) %>% 
  filter(between(preop_psa, 9,11))
