---
title: "Selecting Variables"
output: learnr::tutorial
runtime: shiny_prerendered
---

```{r setup, include=FALSE}
library(learnr)
library(tidyverse)
library(medicaldata)
library(janitor)
library(gradethis)
knitr::opts_chunk$set(echo = FALSE)
prostate <- medicaldata::blood_storage %>% 
  clean_names() %>% as_tibble()
covid <- medicaldata::covid_testing %>% clean_names() %>%  as.tibble()
laryngoscope <- medicaldata::laryngoscope %>% clean_names() %>%  as.tibble()
licorice <- medicaldata::licorice_gargle %>% clean_names() %>%  as.tibble()
```


## Selecting Variables in a dataset

We will start with the prostate dataset, seen here with `glimpse()`:

```{r , echo=TRUE}
glimpse(prostate)
```

### Exercise 1

Write the R code required to select the columns with participant attributes (age, aa, fam_hx) to produce a 3 column dataset.

```{r participant, exercise=TRUE, error=TRUE, lines = 5}
prostate %>% 
  select(---)
```

```{r participant-hint}
prostate %>% 
  select(age, aa, fam_hx)
```

### Exercise 2

Write the R code required to select the variables that contain "recurrence" in the variable name

**Watch** the number of == signs

```{r recur, exercise=TRUE, error=TRUE, lines = 5}
prostate %>% 
  select(---)
```

```{r recur-hint}
prostate %>% 
  select(contains("text"))

```

```{r recur-solution}
prostate %>% 
  select(contains("recurrence"))

```
### Exercise 3

Write the R code required to select all the columns from p_vol to the last column in the prostate dataset (skipping over the first 5).


```{r last, exercise=TRUE, error=TRUE, lines = 5}
prostate %>% 
  select(---)
```

```{r last-hint}
prostate %>% 
  select(p_vol:last_col()) 
```

### Exercise 4

Write the R code required to select the variable age, and all of the variables that start with preop.
```{r in, exercise=TRUE, error=TRUE, lines = 5}
prostate %>% 
  select(---)
```

```{r in-hint}
prostate %>% 
  select(age, starts_with("preop"))
```

### Exercise 5

Take a look at the licorice gargle data, seen here with `glimpse()`

```{r , echo=TRUE}
glimpse(licorice)
```
Write the R code required to select all the variables that end with pain or cough.

```{r licorice, exercise=TRUE, error=TRUE, lines = 5}
licorice %>% 
  select(---)
```

```{r licorice-hint}
licorice %>% 
  select(ends_with(c("pain", "cough")))
```

### Exercise 6

Write the R code required to select all the variables that match the pattern pacu-digit in 0-9-0.

```{r match, exercise=TRUE, error=TRUE, lines = 5}
licorice %>% 
  select(---)
```

```{r match-hint}
licorice %>% 
  select(matches("pacu[---]0"))
```

```{r match-solution}
licorice %>% 
  select(matches("pacu[0-9]0"))
```

Note that [3,9] would also work for this one.


### Exercise 7

Take a look at the covid testing data, seen here with `glimpse()`

```{r covid-glimpse, echo=TRUE}
glimpse(covid)
```

Write the R code required to select only the character variables which also contain the word "name" in the covid dataset.

```{r covid, exercise=TRUE, error=TRUE, lines = 5}
covid %>% 
  select(---)
```

```{r covid-hint}
covid %>% 
  select(where(-- & --))
```

```{r covid-solution}
covid %>% 
  select(where(is.character) & contains("name"))
```
