---
title: "Renaming Variables"
output: learnr::tutorial
runtime: shiny_prerendered
---

```{r setup, include=FALSE}
library(learnr)
library(tidyverse)
library(medicaldata)
library(janitor)
library(gradethis)
knitr::opts_chunk$set(echo = FALSE)
prostate <- medicaldata::blood_storage %>% 
  clean_names() %>% as_tibble()
covid <- medicaldata::covid_testing %>% clean_names() %>%  as.tibble()
laryngoscope <- medicaldata::laryngoscope %>% clean_names() %>%  as.tibble()
licorice <- medicaldata::licorice_gargle %>% clean_names() %>%  as.tibble()
```


## Renaming Variables in a dataset

We will start with the prostate dataset, seen here with `glimpse()`:

```{r , echo=TRUE}
glimpse(prostate)
```

### Exercise 1

Write the R code required to rename the variable fam_hx to fhx. Remember (new_name = old_name) syntax.

```{r rename1, exercise=TRUE, error=TRUE, lines = 5}
prostate %>% 
  rename(--- = ---)
```

```{r rename1-hint}
prostate %>% 
  rename(fhx = fam_hx)
```

### Exercise 2

Write the R code required to rename all of the variables that start with "preop" by replacing "preop" with "PRE", using `replace_with()`, and the `str_replace()` function 

```{r str, exercise=TRUE, error=TRUE, lines = 5}
prostate %>% 
  rename_with(--- , ---)
```

```{r str-hint}
prostate %>% 
  rename_with(--- , ---)

```

```{r str-solution}
prostate %>% 
  rename_with(~str_replace(., "preop", "PRE_"),
              starts_with("preop"))

```

### Exercise 3

Write the R code required to rsystematically rename all of the variables in the blood storage dataset to upper case with the `toupper()` function.

```{r upper, exercise=TRUE, error=TRUE, lines = 5}
medicaldata::blood_storage%>% 
  rename_with(--- , ---)
```

```{r upper-hint}
medicaldata::blood_storage%>% 
  rename_with(~ ---(.), ---())

```

```{r upper-solution}
medicaldata::blood_storage%>% 
  rename_with(~ toupper(.), everything())

```

### Exercise 4

Write the R code required to rsystematically rename all of the variables in the blood storage dataset to lower_snake_case with `janitor::clean_names()`

```{r clean, exercise=TRUE, error=TRUE, lines = 5}
medicaldata::blood_storage%>% 
  janitor::---()
```

```{r clean-hint}
medicaldata::blood_storage%>% 
  janitor::clean_names()

```

