---
title: "Dplyr Tutorial"
output: learnr::tutorial
runtime: shiny_prerendered
---

```{r setup, include=FALSE}
library(learnr)
knitr::opts_chunk$set(echo = FALSE)
library(tidyverse)
library(medicaldata)
library(santoku)
prostate <- medicaldata::blood_storage
```


## Data Wrangling

Please note, <br>
It is **strongly encouraged** to <br>

- google dplyr functions
- phone a friend
- share screens
- ask questions
<br>

### Exercise 

Edit the code below to  

- filter the prostate data by PreopPSA >5
- filter for PVol >50 cc
- select TVol, T.Stage, AA, TimeToRecurrence
- group_by AA
- summarize the mean_tvol, mean_tstage, and mean_time


```{r wrangle_prostate, exercise=TRUE, exercise.lines = 10, error=TRUE}
prostate %>% 
  filter(---) %>% 
  filter(---) %>% 
  select(---) %>% 
  group_by(---) %>% 
  summarize(---)
```


```{r wrangle_prostate-hint}
prostate %>% 
  filter(PreopPSA >5) %>% 
  filter(PVol >50) %>% 
  select(TVol, T.Stage, AA, TimeToRecurrence) %>% 
  group_by(AA) %>% 
  summarize(mean_tvol = mean(TVol, na.rm = T),
            mean_tstage = mean(T.Stage, na.rm = T),
            mean_time = mean(TimeToRecurrence, na.rm = T))
```


### Recoding with _cut_

Recode the variable Age into a new variable, age_decade, using the {base} function, _cut()_. Use the `breaks` argument.

```{r decade, exercise=TRUE, exercise.lines = 5, error=TRUE}

prostate %>% 
  mutate(age_decade = ---) %>% 
  select(Age, age_decade)
```

```{r decade-hint}

prostate %>% 
  mutate(age_decade = cut(Age, breaks = seq(10,100,10))) %>% 
  select(Age, age_decade)
```

### Recoding with {santoku} _chop_

Google the R package {santoku}, and the _chop()_ function (with the breaks argument). Use this to create age_decade from the Age variable.

```{r chop, exercise=TRUE, exercise.lines = 5, error=TRUE}
prostate %>% 
  mutate(age_decade = --- ) %>% 
  select(Age, age_decade)
```


```{r chop-hint}
prostate %>% 
  mutate(age_decade = chop(Age, breaks = c(seq(10,100, 10)))) %>% 
  select(Age, age_decade)
```

